Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated Model 1 section to align with structure of model 2 and 3

Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of suscesptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018). Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls. Alternatively, the likelihood of reduced mobility for those living under the poverty line and/or who are unemployed may also play a role in the relationship with case rate. As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.

@@

From the above plot, there is noticeable variation in the smoothed blue line. However, it can be posited that there is a inverse relationship between the variables given the roughly linear relationship in the overall downward trend in the line. In order to test this, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.

@@

To improve the precision of the t-test of coefficients, classical standard errors can be used over robust standard errors to test the regression. To determine the applicability of using classical standard errors, in addition to the three Classical Linear Model (CLM) assumptions necessary for the robust standard errors, the CLM assumption for homoskedastic conditional errors should be evaluated. If this additional assumption is met, the robust standard errors can be replaced by the classical standard errors. The assumption of homoskedasticity is met if the data does not have large variance among the residuals.

To evaluate this assumption, the square root of the residuals (to remove any negative values) is plotted against the fitted values of the model. Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.

@@

Next, the other CLM assumptions are revisited to determine the whether the coefficients are unbiased and consistent. However, the assumption for IID is not detailed here given that it is evaluated more generally for all three models within section 3.

The second CLM assumption to be evaluated here is that of a linear conditional expectation relationship within the data, so that the model is accurately estimating the relationship between the variables. To test for this, the residuals of model 1 are plotted against its fitted values.

The third CLM assumption examined for model 1 is the assumption of normality of errors. This is a necessary assumption to ascertain that the errors used in the regression are drawn from a normal distribution, so that they can be accepted when used to calculate the significance levels. Both a QQ plot and histogram of the distribution of the residuals is plotted to observe the normality.

@@

Finally the Cook's distance is examined by a plot of the residuals vs the model 1 leverage to look for any outliers in the values. From the plot below, no obvious outliers are detected.

@@

It is also noted that there is no test conducted for collinearity in model 1 given there is only one input variable.

>>>>

Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018). Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls. Alternatively, the likelihood of reduced mobility for those living under the poverty line and/or who are unemployed may also play a role in the relationship with case rate. As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.

@@

From the above plot, there is noticeable variation in the smoothed blue line. However, it can be posited that there is a inverse relationship between the variables given the roughly linear relationship in the overall downward trend in the line.

#### Model 1 Regression

In order to test the relationship between the variables, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.

@@

To improve the precision of the t-test of coefficients, classical standard errors can be used over robust standard errors to test the regression. To determine the applicability of using classical standard errors, in addition to the three Classical Linear Model (CLM) assumptions necessary for the robust standard errors, the CLM assumption for homoskedastic conditional errors should be evaluated. If this additional assumption is met, the robust standard errors can be replaced by the classical standard errors. The assumption of homoskedasticity is met if the data does not have large variance among the residuals. The analysis for this assumption, as well as the other CLM assumptions, is conducted in the following subsection titled "Model 1 CLM Assumptions".

Below, the t-test of coefficients is run on the regression model above.

#### Model 1 CLM Assumptions

##### i. IID Sampling

The first assumption of IID is not detailed within this section given that it is evaluated more generally for all three models within section 3.

##### ii. Linear Conditional Expectation

The second CLM assumption to be evaluated is that of a linear conditional expectation relationship within the data, so that the model is accurately estimating the relationship between the variables. To test for this, the residuals of model 1 are plotted against its fitted values.

##### iii. No Perfect Collinearity

There is no test conducted for the third CLM assumption of no perfect collinearity in model 1 given there is only one input variable.

##### iv. Homoskedastic Errors

To evaluate the fourth CLM assumption of homoskedastic errors, the square root of the residuals (to remove any negative values) is plotted against the fitted values of the model. Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.

@@

##### v. Normally Distributed Errors

The final CLM assumption examined for model 1 is the assumption of normality of errors. This is a necessary assumption to ascertain that the errors used in the regression are drawn from a normal distribution, so that they can be accepted when used to calculate the significance levels. Both a QQ plot and histogram of the distribution of the residuals is plotted to observe the normality.

@@

#### Influence of Data Points in Model 1

In addition to the CLM assumptions, the Cook's distance is also examined by a plot of the residuals vs the model 1 leverage to estimate the influence of the data points and identify any outliers in the values. From the plot below, no obvious outliers are detected.

@@
